---
title: "Final Project"
author: "Jinghan Fu Dongyu Zhou Rebecca Brent Kepan Gao"
date: "2020/3/30"
output:
  pdf_document: default
---

#Loading and Exploring Data__
Loading R packages used besides base R.
```{r, message=FALSE, warning=FALSE,echo=FALSE}
library(knitr)
library(ggplot2)
library(plyr)
library(dplyr)
library(corrplot)
library(caret)
library(gridExtra)
library(scales)
library(Rmisc)
library(ggrepel)
library(randomForest)
library(psych)
library(xgboost)
library(PerformanceAnalytics)
library(naniar)
library(glmnet)
library(leaps)
library(pls)
library(splines)
library(gam)
library(tree)
library(gbm)
```

```{r}
train <- read.csv("~/DS5110/house-prices-advanced-regression-techniques/train.csv", stringsAsFactors = F)
test <- read.csv("~/DS5110/house-prices-advanced-regression-techniques/test.csv", stringsAsFactors = F)
```


```{r}
dim(train)
str(train)
res <- sapply(train,class)
table(res)
```

#Data pre-processing
```{r}
Id_labels <- test$Id
test$Id <- NULL
train$Id <-NULL
test$SalePrice <- NA
```


```{r}
data <- rbind(train, test)
dim(data)
```

#Variable Exploration & Missing data imputation

```{r, message=FALSE}
ggplot(data=data[!is.na(data$SalePrice),], aes(x=SalePrice)) +
        geom_histogram(fill="darkcyan") +
        ggtitle("Right skewd distribution")
```

```{r}
summary(data$SalePrice)
```
*Righr skewd, because most of people cannot affort high-end luxury house.*

```{r}
numeric.var <- which(sapply(data, is.numeric)) 
length(numeric.var)
numeric.name <- names(numeric.var) 
#numeric.name
```


```{r}
data.numeric <- data[, numeric.var]
cor.numvar <- cor(data.numeric, use="pairwise.complete.obs", method = "pearson") 
#cor.numvar
cor.sorted <- as.matrix(sort(cor.numvar[,'SalePrice'], decreasing = T))
high.cor <- names(which(apply(cor.sorted, 1 , function(x) abs(x)>0.5)))
#high.cor
cor.numvar <- cor.numvar[high.cor, high.cor]
heatmap(cor.numvar, Colv = NA, Rowv = NA, scale="column")
corrplot.mixed(cor.numvar,tl.col="chocolate2", tl.pos = "lt",tl.cex = 0.9,cl.cex = 0.9, number.cex=0.9,outline= TRUE)
```
We would like to deal with the top two variable since their correlation is > 0.7
```{r}
ggplot(data=data[!is.na(data$SalePrice),], aes(x=factor(OverallQual), y=SalePrice))+
        geom_boxplot(col='darkolivegreen') + labs(x='Overall material and finish quality') 
```
*Outlier may exist in grade 4. We can check more variables to find out.*
```{r}
ggplot(data=data[!is.na(data$SalePrice),], aes(x=GrLivArea, y=SalePrice))+
        geom_point(size=2, shape=23,col='coral2')+
        labs(x='Above grade (ground) living area square feet')+
        geom_smooth(method = "lm",color="coral4")+
        geom_text_repel(aes(label = ifelse(train$GrLivArea[!is.na(train$SalePrice)]>4000, rownames(data), '')))
```

```{r}
data[c(524,692,1183,1299), c('SalePrice', 'GrLivArea', 'OverallQual')]
```
*524 and 1299 are pretty odd, their overallQual is 10 but their saleprice are way too low. We can consider them as outlier and remove it later.*

We would like to see which variables contain missing values.
```{r}
gg_miss_upset(data,nset = 8)
```

```{r}
NAcol <- sapply(data,function(x) sum(is.na(x)))
miss <- sort(NAcol,decreasing = T)
miss[miss>0]
length(miss[miss>0])-1
```

*The Na of SalePrice is from the test dataset, so PoolQC, MiscFeature, Alley, Fence, FireplaceQu are top 5 variables with missing value. Based on the data description, we can deal with variable by group, such as pool, basement etc,.*

```{r}
for(x in c("PoolQC","MiscFeature","Alley","Fence","FireplaceQu")){
        data[is.na(data[,x]),x]<- "None"
}
```

##Pool Varioble
```{r}
#data$PoolArea
```
*Most of the Pool Area is 0 so those NAs in the PoolQC make sense, so we impute 0 for NAs.*
```{r}
table(data$PoolQC)
Qualities <- c('None' = 0, 'Po' = 1, 'Fa' = 2, 'TA' = 3, 'Gd' = 4, 'Ex' = 5)
data$PoolQC<-as.integer(revalue(data$PoolQC, Qualities))
```

```{r}
ggplot(data=data[!is.na(data$SalePrice),], aes(x=factor(PoolQC), y=SalePrice))+
        geom_boxplot(col='darkcyan') + labs(x='Pool quality') 
```
*House with higher pool quality have higehr price as we predict but there are also some high house price when those houses does not have pools.*

```{r}
data[data$PoolArea>0 & data$PoolQC==0, c('PoolArea', 'PoolQC', 'OverallQual')]
```
*Imputing NAs in the PoolQC based on the average quality of OverallQual *
```{r}
data$PoolQC[2421] <- 2
data$PoolQC[2504] <- 3
data$PoolQC[2600] <- 2
```


##Miscellaneous Feature variables
```{r}
table(data$MiscFeature)
data$MiscFeature <- as.factor(data$MiscFeature)
```
```{r}
ggplot(data=data[!is.na(data$SalePrice),], aes(x=factor(MiscFeature), y=SalePrice))+
  geom_bar(stat = 'summary',fun.y='median',fill='darkolivegreen4') + 
  labs(x='Miscellaneous feature ') 
```

```{r}
data[data$MiscVal!=0 & data$MiscFeature=="None", c('MiscVal', 'MiscFeature', 'OverallQual')]
```

```{r}
data$MiscFeature[2550] <- 'Shed'
```

##Alley

```{r}
table(data$Alley)
data$Alley <- as.factor(data$Alley)
```
```{r}
ggplot(data=data[!is.na(data$SalePrice),], aes(x=factor(Alley), y=SalePrice))+
  geom_bar(stat = 'summary',fun.y='median',fill='cornflowerblue')+
  geom_label(stat = 'count',aes(label=..count..,y=..count..)) + 
  labs(x='Type of alley access to property')

```
*House with paved alley seems to have higher price*

##Fence
```{r}
table(data$Fence)
data$Fence <- as.factor(data$Fence)
```
```{r}
ggplot(data=data[!is.na(data$SalePrice),], aes(x=factor(Fence), y=SalePrice))+
  geom_bar(stat = 'summary',fun.y='median',fill='aquamarine4')+
  geom_label(stat = 'count',aes(label=..count..,y=..count..)) + 
  labs(x='Fence quality')

```
*GoodPrv seems to have higher house price*

##Fireplace Variable
```{r}
table(data$FireplaceQu)
data$FireplaceQu<-as.integer(revalue(data$FireplaceQu, Qualities))
table(data$Fireplaces)
```
```{r}
ggplot(data=data[!is.na(data$SalePrice),], aes(x=factor(FireplaceQu), y=SalePrice))+
  geom_bar(stat = 'summary',fun.y='median',fill='aquamarine4')+
  geom_label(stat = 'count',aes(label=..count..,y=..count..)) + 
  labs(x='Fireplace quality')
```
*House with higher Fireplace Quality usually have higher price*


##Lot Variables
```{r}
#table(data$LotFrontage)
```

```{r}
#table(data$LotFrontage)
for (i in 1:nrow(data)){
        if(is.na(data$LotFrontage[i])){
               data$LotFrontage[i] <- as.integer(median(data$LotFrontage[data$Neighborhood==data$Neighborhood[i]], na.rm=TRUE)) 
        }
}
any(is.na(data$LotFrontage))

#data$LotShape
table(data$LotShape)
data$LotShape<-as.integer(revalue(data$LotShape, c('IR3'=0, 'IR2'=1, 'IR1'=2, 'Reg'=3)))
#table(data$LotArea)

table(data$LotConfig)
data$LotConfig <- as.factor(data$LotConfig)
```

##Garage Variables
GarageYrBlt GarageFinish   GarageQual   GarageCond   GarageType    
    159          159          159          159          157       
*We assume that the first four should have the same observation NAs, GarageType has 157 NAs, there might be lesser houses without a Garage.*
```{r}
table(data$GarageYrBlt)
#data[data$GarageYrBlt==2207,c("GarageYrBlt","YearBuilt")]
data$GarageYrBlt[2593]<-2007 #typo
```

```{r}
data$GarageYrBlt[is.na(data$GarageYrBlt)] <- data$YearBuilt[is.na(data$GarageYrBlt)]
any(is.na(data$GarageYrBlt))
data[!is.na(data$GarageType) & is.na(data$GarageFinish), c('GarageCars', 'GarageArea', 'GarageType', 'GarageCond', 'GarageQual', 'GarageFinish')]
```

```{r}
names(sort(table(data$GarageCond),decreasing = T))[1]
data$GarageCond[2127] <-  "TA"

names(sort(table(data$GarageQual),decreasing = T))[1]
data$GarageQual[2127] <-  "TA"

names(sort(table(data$GarageFinish),decreasing = T))[1]
data$GarageFinish[2127] <-"Unf"

data$GarageCars[2577] <- 0
data$GarageArea[2577] <- 0
data$GarageType[2577] <- NA
```
*Imputing mode into the house 2123, as for house 2577, just using none and 0 for those garage varible since most of the variable is NA*

```{r}
for(gar in c("GarageFinish","GarageQual","GarageCond")){
        data[is.na(data[,gar]),gar]<-"None"
}
```

```{r}
table(data$GarageType)
data$GarageType[is.na(data$GarageType)] <- 'No Garage'
data$GarageType <- as.factor(data$GarageType)

table(data$GarageFinish)
Finish <- c('None'=0, 'Unf'=1, 'RFn'=2, 'Fin'=3)
data$GarageFinish<-as.integer(revalue(data$GarageFinish, Finish))

table(data$GarageQual)
data$GarageQual<-as.integer(revalue(data$GarageQual, Qualities))

table(data$GarageCond)
data$GarageCond<-as.integer(revalue(data$GarageCond, Qualities))

#data$GarageArea
#data$GarageCars
table(data$GarageCars)
```
##Basement Varibales
BsmtCond BsmtExposure BsmtQual BsmtFinType2 BsmtFinType1 BsmtFullBath BsmtHalfBath 
82           82         81           80           79      2            2  
BsmtFinSF1   BsmtFinSF2 BsmtUnfSF TotalBsmtSF 
         1            1         1           1

```{r}
data[!is.na(data$BsmtFinType1) & (is.na(data$BsmtCond)|is.na(data$BsmtQual)|is.na(data$BsmtExposure)|is.na(data$BsmtFinType2)), c('BsmtCond','BsmtExposure','BsmtQual','BsmtFinType2', 'BsmtFinType1')]
```

```{r}
names(sort(table(data$BsmtCond),decreasing = T))[1]
data$BsmtCond[c(2041, 2186, 2525)] <- "TA"
names(sort(table(data$BsmtExposure),decreasing = T))[1]
data$BsmtExposure[c(949, 1488, 2349)] <- "No"
names(sort(table(data$BsmtQual),decreasing = T))[1]
data$BsmtQual[c(2218, 2219)] <- "TA"
names(sort(table(data$BsmtFinType2),decreasing = T))[1]
data$BsmtFinType2[333] <- "Unf"
```

```{r}
for(x in c("BsmtCond","BsmtExposure","BsmtQual","BsmtFinType2","BsmtFinType1")){
      data[is.na(data[,x]),x]<-"None"
}
```

```{r}
data$BsmtQual[is.na(data$BsmtQual)] <- 'None'
table(data$BsmtCond)
data$BsmtCond<-as.integer(revalue(data$BsmtCond, Qualities))

table(data$BsmtExposure)
Exposure <- c('None'=0, 'No'=1, 'Mn'=2, 'Av'=3, 'Gd'=4)
data$BsmtExposure<-as.integer(revalue(data$BsmtExposure, Exposure))

table(data$BsmtQual)
data$BsmtQual<-as.integer(revalue(data$BsmtQual, Qualities))

table(data$BsmtFinType2)
FinType <- c('None'=0, 'Unf'=1, 'LwQ'=2, 'Rec'=3, 'BLQ'=4, 'ALQ'=5, 'GLQ'=6)
data$BsmtFinType2<-as.integer(revalue(data$BsmtFinType2, FinType))

table(data$BsmtFinType1)
data$BsmtFinType1<-as.integer(revalue(data$BsmtFinType1, FinType))
```
BsmtFullBath BsmtHalfBath  BsmtFinSF1   BsmtFinSF2 BsmtUnfSF TotalBsmtSF 
       2            2            1            1         1           1
```{r}
data[(is.na(data$BsmtFullBath)|is.na(data$BsmtHalfBath)|is.na(data$BsmtFinSF1)|is.na(data$BsmtFinSF2)|is.na(data$BsmtUnfSF)|is.na(data$TotalBsmtSF)), c("BsmtQual","BsmtFullBath", "BsmtHalfBath","BsmtFinSF1", "BsmtFinSF2", "BsmtUnfSF", "TotalBsmtSF" )]
```


```{r}
table(data$BsmtFullBath)
data$BsmtFullBath[is.na(data$BsmtFullBath)] <-0
table(data$BsmtHalfBath)
data$BsmtHalfBath[is.na(data$BsmtHalfBath)] <-0
#table(data$BsmtFinSF1)
data$BsmtFinSF1[is.na(data$BsmtFinSF1)] <-0
#table(data$BsmtFinSF2)
data$BsmtFinSF2[is.na(data$BsmtFinSF2)] <-0
#table(data$BsmtUnfSF)
data$BsmtUnfSF[is.na(data$BsmtUnfSF)] <-0
#table(data$TotalBsmtSF)
data$TotalBsmtSF[is.na(data$TotalBsmtSF)] <-0
```
##Masonry Variables

MasVnrType   MasVnrArea
        24           23
```{r}
data[!is.na(data$MasVnrArea) & is.na(data$MasVnrType), c('MasVnrType', 'MasVnrArea')]
```

```{r}
names(sort(table(data$MasVnrType),decreasing = T))[2]
data$MasVnrType[2611] <- "BrkFace"
data[2611, c('MasVnrType', 'MasVnrArea')]
data$MasVnrType[is.na(data$MasVnrType)] <- 'None'
Masonry <- c('None'=0, 'BrkCmn'=0, 'BrkFace'=1, 'Stone'=2)
data$MasVnrType<-as.integer(revalue(data$MasVnrType, Masonry))

table(data$MasVnrType)
data$MasVnrArea[is.na(data$MasVnrArea)] <-0
```
```{r}
ggplot(data[!is.na(data$SalePrice),],aes(x=MasVnrType,y=SalePrice))+
geom_bar(stat = 'summary',fun.y='median',fill='plum3')+
geom_label(stat='count',aes(label=..count..,y=..count..))
```



```{r}
table(data$MSZoning)
names(sort(table(data$MSZoning),decreasing = T))[1]
data$MSZoning[is.na(data$MSZoning)] <- "RL"
data$MSZoning <- as.factor(data$MSZoning)
```
__Utilities & Functional__
```{r}
table(data$Utilities)
data$Utilities <- NULL
```
*Since Most of the Utilitis are AllPub, we can ignore this Variable*

```{r}
table(data$Functional)
names(sort(table(data$Functional),decreasing = T))[1]
data$Functional[is.na(data$Functional)] <- "Typ"
Functions<-c('Sal'=0, 'Sev'=1, 'Maj2'=2, 'Maj1'=3, 'Mod'=4, 'Min2'=5, 'Min1'=6, 'Typ'=7)
data$Functional <- as.integer(revalue(data$Functional,Functions))
```

```{r}
table(data$Electrical)
names(sort(table(data$Electrical),decreasing = T))[1]
data$Electrical[is.na(data$Electrical)] <-"SBrkr"
data$Electrical <- as.factor(data$Electrical)
```

```{r}
table(data$KitchenQual)
names(sort(table(data$KitchenQual),decreasing = T))[1]
data$KitchenQual[is.na(data$KitchenQual)] <- 'TA'
data$KitchenQual<-as.integer(revalue(data$KitchenQual, Qualities))
table(data$KitchenAbvGr)
```
##Exterior Variable
Exterior1st  Exterior2nd   
          1            1  

```{r}
table(data$Exterior1st)
names(sort(table(data$Exterior1st),decreasing = T))[1]
data$Exterior1st[is.na(data$Exterior1st)] <-"VinylSd"
data$Exterior1st <- as.factor(data$Exterior1st)

table(data$Exterior2nd)
names(sort(table(data$Exterior2nd),decreasing = T))[1]
data$Exterior2nd[is.na(data$Exterior2nd)] <- "VinylSd"
data$Exterior2nd <- as.factor(data$Exterior2nd)

table(data$ExterQual)
data$ExterQual<-as.integer(revalue(data$ExterQual, Qualities))

table(data$ExterCond)
data$ExterCond<-as.integer(revalue(data$ExterCond, Qualities))
```
##Sale Variables
```{r}
table(data$SaleType)
names(sort(table(data$SaleType),decreasing = T))[1]
data$SaleType[is.na(data$SaleType)] <- "WD"
data$SaleType <- as.factor(data$SaleType)

table(data$SaleCondition)
data$SaleCondition <- as.factor(data$SaleCondition)
```

##Categorical variables
```{r}
cha.var <- which(sapply(data, is.character)) 
cha.name <- names(cha.var)
cha.name
```
 
```{r}
table(data$Street)
data$Street<-as.integer(revalue(data$Street, c('Grvl'=0, 'Pave'=1)))
```
```{r}
table(data$LandContour)
data$LandContour <- as.factor(data$LandContour)
```
```{r}
table(data$LandSlope)
data$LandSlope<-as.integer(revalue(data$LandSlope, c('Sev'=0, 'Mod'=1, 'Gtl'=2)))
```
```{r}
table(data$Neighborhood)
data$Neighborhood <- as.factor(data$Neighborhood)
```
```{r}
table(data$Condition1)
data$Condition1 <- as.factor(data$Condition1)
```
```{r}
table(data$Condition2)
data$Condition2 <- as.factor(data$Condition2)
```
```{r}
table(data$BldgType)
data$BldgType <- as.factor(data$BldgType)

```

```{r}
table(data$HouseStyle)  
data$HouseStyle <- as.factor(data$HouseStyle)

```

```{r}
table(data$RoofStyle)
data$RoofStyle <- as.factor(data$RoofStyle)

```


```{r}
table(data$CentralAir)
data$CentralAir<-as.integer(revalue(data$CentralAir, c('N'=0, 'Y'=1)))

```

```{r}
table(data$HeatingQC)
data$HeatingQC<-as.integer(revalue(data$HeatingQC, Qualities))

```
```{r}
table(data$PavedDrive)
data$PavedDrive<-as.integer(revalue(data$PavedDrive, c('N'=0, 'P'=1, 'Y'=2)))

```

```{r}
table(data$Heating)
data$Heating <- as.factor(data$Heating)
```

```{r}
table(data$Foundation)
data$Foundation <- as.factor(data$Foundation)
```

```{r}
table(data$RoofMatl)
data$RoofMatl <- as.factor(data$RoofMatl)
```


```{r}
table(data$MoSold)
data$MoSold <- as.factor(data$MoSold)
```
```{r}
table(data$MSSubClass)
data$MSSubClass <- as.factor(data$MSSubClass)
```

##Important variables

```{r}
data <-data %>% mutate("Remod" = ifelse(data$YearBuilt==data$YearRemodAdd, 0, 1),
                       "Age" = as.numeric(data$YrSold)-data$YearRemodAdd,
                       "New" = ifelse(data$YrSold==data$YearBuilt, 1, 0), 
                       "TotalSF" = TotalBsmtSF + X1stFlrSF+ X2ndFlrSF + GarageArea, 
                       "PropGrLivArea" = GrLivArea / TotalSF,
                       "TotalEstLivableSF" = TotalSF - BsmtUnfSF - GarageArea,
                        "PropLivGrLivArea" = GrLivArea / TotalEstLivableSF,
                       "PropUnlivArea" = (BsmtUnfSF + GarageArea) / TotalSF,
                       "TotalFullBath" = BsmtFullBath + FullBath,
                       "TotalHalfBath" = BsmtHalfBath + HalfBath,
                       "TotalBath" = TotalFullBath + .5*TotalHalfBath, 
                       "TotRooms" = TotRmsAbvGrd + TotalBath,
                       "HasFence" = ifelse(Fence=='None', 0, 1),
                       "HasPool" = ifelse(PoolArea > 0, 1, 0),
                       "HasWoodDeck" = ifelse(WoodDeckSF > 0, 1,0),
                       "HasPorch" = ifelse(OpenPorchSF > 0 | EnclosedPorch > 0| X3SsnPorch > 0 | ScreenPorch > 0, 1, 0),
                       "HasGarage" = ifelse(GarageType=='No Garage', 0, 1),
                       "HasAlleyAccess" = ifelse(Alley=='None', 0, 1),
                       "HasOpenPorch" = ifelse(OpenPorchSF >0, 1, 0),
                       "HasEnclosedPorch" = ifelse(EnclosedPorch > 0, 1, 0),
                       "HasX3SsnPorch" = ifelse(X3SsnPorch > 0, 1 , 0),
                       "HasScreenPorch" = ifelse(ScreenPorch > 0, 1, 0),
                       "HasBsmt" = ifelse(TotalBsmtSF > 0, 1, 0))

#Any houses that are less than 2 Stories should not have 2nd floor square footage. Any 2nfFlrSF for 1.5 story houses will be added to the 1stFlrSF and 2ndFlrSF will be set to 0. X2ndFlr will be set to 'N' for these houses.
data <- data %>%
  mutate("X1stFlrSF" = ifelse(HouseStyle == '1.5Fin' | HouseStyle == '1.5Unf'| HouseStyle == '1Story', X1stFlrSF+X2ndFlrSF, X1stFlrSF),
         "X2ndFlrSF" = ifelse(HouseStyle == '1.5Fin' | HouseStyle == '1.5Unf' | HouseStyle == '1Story', 0, X2ndFlrSF),
         "X2ndFlr" = ifelse(X2ndFlrSF > 0,1,0))


data$YrSold <- as.factor(data$YrSold)
data$Remod <-as.factor(data$Remod)
data$New <-as.factor(data$New)
data$HasFence <- as.factor(data$HasFence)
data$HasPool <- as.factor(data$HasPool)
data$HasPorch <- as.factor(data$HasPorch)
data$HasGarage <- as.factor(data$HasGarage)
data$HasAlleyAccess <- as.factor(data$HasAlleyAccess)
data$HasOpenPorch <- as.factor(data$HasOpenPorch)
data$HasEnclosedPorch <- as.factor(data$HasEnclosedPorch)
data$HasX3SsnPorch <- as.factor(data$HasX3SsnPorch)
data$HasScreenPorch <- as.factor(data$HasScreenPorch)
data$HasBsmt <- as.factor(data$HasBsmt)
data$X2ndFlr <- as.factor(data$X2ndFlr)


data$ConditionNormal <- factor(with(data, ifelse(Condition1=='Norm' | Condition2=='Norm', 1, 0)))

data$ConditionArtery <- factor(with(data, ifelse(Condition1=='Artery'|Condition2=='Artery', 1, 0)))

data$ConditionPosA <- factor(with(data, ifelse(Condition1=='PosA'|Condition2=='PosA', 1, 0)))

data$ConditionPosN <- factor(with(data, ifelse(Condition1=='PosN'|Condition2=='PosN', 1, 0)))

data$ConditionRRAe <- factor(with(data, ifelse(Condition1=='RRAe'|Condition2=='RRAe', 1, 0)))

data$ConditionRRAn <- factor(with(data, ifelse(Condition1=='RRAn'|Condition2=='RRAn', 1, 0)))

data$ConditionRRNe <- factor(with(data, ifelse(Condition1=='RRNe'|Condition2=='RRNe', 1, 0)))

data$ConditionRRNn <- factor(with(data, ifelse(Condition1=='RRNn'|Condition2=='RRNn', 1, 0)))

data$ExteriorAsbShng <- factor(with(data, ifelse(Exterior1st=='AsbShng'|Exterior2nd=='AsbShng', 1, 0)))

data$ExteriorAsphShn <- factor(with(data, ifelse(Exterior1st=='AsphShn'|Exterior2nd=='AsphShn', 1, 0)))

data$ExteriorBrkComm <- factor(with(data, ifelse(Exterior1st=='BrkComm'|Exterior2nd=='BrkComm', 1, 0)))

data$ExteriorBrkFace <- factor(with(data, ifelse(Exterior1st=='BrkFace'|Exterior2nd=='BrkFace', 1, 0)))

data$ExteriorCBlock <- factor(with(data, ifelse(Exterior1st=='CBlock'|Exterior2nd=='CBlock', 1, 0)))

data$ExteriorCmentBd <- factor(with(data, ifelse(Exterior1st=='CmentBd'|Exterior2nd=='CmentBd', 1, 0)))

data$ExteriorHdBoard <- factor(with(data, ifelse(Exterior1st=='HdBoard'|Exterior2nd=='HdBoard', 1, 0)))

data$ExteriorImStucc <- factor(with(data, ifelse(Exterior1st=='ImStucc'|Exterior2nd=='ImStucc', 1, 0)))

data$ExteriorMetalSd <- factor(with(data, ifelse(Exterior1st=='MetalSd'|Exterior2nd=='MetalSd', 1, 0)))

data$ExteriorOther <- factor(with(data, ifelse(Exterior1st=='Other'|Exterior2nd=='Other', 1, 0)))

data$ExteriorPlywood <- factor(with(data, ifelse(Exterior1st=='Plywood'|Exterior2nd=='Plywood', 1, 0)))

data$ExteriorStone <- factor(with(data, ifelse(Exterior1st=='Stone'|Exterior2nd=='Stone', 1, 0)))

data$ExteriorStucco <- factor(with(data, ifelse(Exterior1st=='Stucco'|Exterior2nd=='Stucco', 1, 0)))

data$ExteriorVinylSd <- factor(with(data, ifelse(Exterior1st=='VinylSd'|Exterior2nd=='VinylSd', 1, 0)))

data$ExteriorWdSdng <- factor(with(data, ifelse(Exterior1st=='Wd Sdng'|Exterior2nd=='Wd Sdng', 1, 0)))

data$ExteriorWdShng <- factor(with(data, ifelse(Exterior1st=='Wd Shng'|Exterior2nd=='Wd Shng', 1, 0)))
```


```{r}
numeric.var <- which(sapply(data, is.numeric)) 
factor.var <- which(sapply(data, is.factor)) 
names(factor.var)
names(numeric.var)
length(numeric.var)
length(factor.var)
```

```{r}
data.numeric <- data[, numeric.var]
cor.numvar <- cor(data.numeric, use="pairwise.complete.obs", method = "pearson") 
#cor.numvar
cor.sorted <- as.matrix(sort(cor.numvar[,'SalePrice'], decreasing = T))
high.cor <- names(which(apply(cor.sorted, 1 , function(x) abs(x)>0.5)))
#high.cor
cor.numvar <- cor.numvar[high.cor, high.cor]
heatmap(cor.numvar, Colv = NA, Rowv = NA, scale="column")
corrplot.mixed(cor.numvar, tl.col="chocolate2", tl.pos = "lt",tl.cex = 0.6,cl.cex = 0.6, number.cex=0.6,outline= TRUE)
```

```{r}
drop.var <- c('YearRemodAdd', 'GarageYrBlt', 'GarageArea', 'GarageCond', 'TotalRmsAbvGrd', 'BsmtFinSF1','X1stFlrSF', 'Condition1', 'Condition2', 'Exterior1st', 'Exterior2nd')

data <- data[,!(names(data) %in% drop.var)]
```

```{r}
data <- data[-c(524, 1299),]
```



```{r}
numeric.name <- numeric.name[!(numeric.name %in% c('MSSubClass', 'MoSold', 'YrSold', 'SalePrice', 'OverdataQual', 'OverdataCond'))] 
numeric.name <- append(numeric.name, c('Age', 'TotalSqFeet'))

DFnumeric <- data[,names(data) %in% numeric.name]
DFfactors <- data[,!(names(data) %in% numeric.name)]
DFfactors <- DFfactors[, names(DFfactors) != 'SalePrice']
length(DFnumeric)
length(DFfactors)
```
```{r}
for(i in 1:ncol(DFnumeric)){
        if (abs(skew(DFnumeric[,i]))>0.8){
                DFnumeric[,i] <- log(DFnumeric[,i] +1)
        }
}
```
```{r}
PreNum <- preProcess(DFnumeric, method=c("center", "scale"))
#PreNum
DFnorm <- predict(PreNum, DFnumeric)
dim(DFnorm)
DFdummies <- as.data.frame(model.matrix(~.-1, DFfactors))
dim(DFdummies)
```

```{r}
data1 <- cbind(DFnorm, DFdummies)
```


```{r}
data$SalePrice <- log(data$SalePrice) 
skew(data$SalePrice)
qqnorm(data$SalePrice)
qqline(data$SalePrice)
```

```{r}
ggplot(data=data[!is.na(data$SalePrice),], aes(x=SalePrice)) +
        geom_histogram(fill="darkcyan") +
        ggtitle("Lognormal distribution")
```


```{r}
train <- data1[!is.na(data$SalePrice),]
test <- data1[is.na(data$SalePrice),]
```

#Model Building
##Linear Regression
```{r}
lm.base<-lm(data$SalePrice[!is.na(data$SalePrice)] ~.,train)
par(mfrow = c(2, 2))
plot(lm.base)
```
```{r}
lm.pred <- predict(lm.base,test)
pred<-exp(lm.pred)
prediction1 <- data.frame(Id = Id_labels,SalePrice = pred)
write.csv(prediction1, file = 'prediction1.csv', row.names = F)
```
*0.13048 from Kaggle*


##Stepwise Selection
###Forward selection
```{r}
empty <- lm(data$SalePrice[!is.na(data$SalePrice)] ~1, data=train)
full <- lm(data$SalePrice[!is.na(data$SalePrice)] ~., data=train) 
set.seed(2)
lm.for <- step(empty, scope=list(lower=empty, upper=full), direction="forward")
summary(lm.for)
lm.pred <- predict(lm.for,test)
res1 <- data.frame(Id = Id_labels, SalePrice = exp(lm.pred))
write.csv(res1, file = "forward.csv", row.names = FALSE)
```
*score 0.12583 From Kaggle*

###Both selection
```{r}
empty <- lm(data$SalePrice[!is.na(data$SalePrice)] ~1, data=train)
full <- lm(data$SalePrice[!is.na(data$SalePrice)] ~., data=train) 
set.seed(2)
lm.for <- step(empty, scope=list(lower=empty, upper=full), direction="both")
summary(lm.for)
lm.pred <- predict(lm.for,test)
both <- data.frame(Id = Id_labels, SalePrice = exp(lm.pred))
write.csv(both, file = "both.csv", row.names = FALSE)
```
*.012623 from Kaggle*
##Shrinkage Methods
###Ridge
```{r}
X<- model.matrix(data$SalePrice[!is.na(data$SalePrice)] ~.,data = train)[,-1]
y<- data$SalePrice[!is.na(data$SalePrice)]
set.seed(101)
ridge.model<-glmnet(X,y,alpha = 0)
ridge.model$lambda
coef(ridge.model, s = ridge.model$lambda[100])
coef(ridge.model, s = ridge.model$lambda[1])
cv.out <- cv.glmnet(X,y,alpha = 0, nfolds = 10)
cv.out$lambda.min
cv.out$cvm
plot(cv.out)
cv.out <- cv.glmnet(X,y,alpha = 1, nfolds = 10)
ridge_coef<-predict(cv.out, type = "coefficients", s = cv.out$lambda.min)
length(ridge_coef[ridge_coef != 0]) #variable used
length(ridge_coef[ridge_coef == 0]) #variable not used
```
```{r}
test.X <- as.matrix(test)
pred<-predict(ridge.model, s = cv.out$lambda.min, newx = test.X)[,1]
predictions<-exp(pred)
prediction_ridge<-exp(pred)
prediction3 <- data.frame(Id = Id_labels,SalePrice = predictions)
#prediction3
write.csv(prediction3, file = 'prediction3.csv', row.names = F)
```
*0.12337 from kaggle*

###Lasso:
```{r}
X<- model.matrix(data$SalePrice[!is.na(data$SalePrice)] ~.,data = train)[,-1]
y<- data$SalePrice[!is.na(data$SalePrice)]
set.seed(101)
lasso.model<-glmnet(X,y,alpha = 1,lambda = seq(0.001,0.1,by = 0.0005))
lasso.model$lambda
cv.out <- cv.glmnet(X,y,alpha = 1, nfolds = 10)
cv.out$lambda.min
cv.out$cvm
plot(cv.out)
lasso_coef<-predict(cv.out, type = "coefficients", s = cv.out$lambda.min)
length(lasso_coef[lasso_coef != 0]) #variable used
length(lasso_coef[lasso_coef == 0]) #variable not used
```
```{r}
test.X <- as.matrix(test)
pred<-predict(lasso.model, s = cv.out$lambda.min, newx = test.X)[,1]
predictions<-exp(pred)
prediction_lasso<-exp(pred)
prediction4 <- data.frame(Id = Id_labels,SalePrice = predictions)
#prediction4
write.csv(prediction4, file = 'prediction4.csv', row.names = F)
```
*0.12078 from kaggle*

###Elastic Net
```{r}
X<- model.matrix(data$SalePrice[!is.na(data$SalePrice)] ~.,data = train)[,-1]
y<- data$SalePrice[!is.na(data$SalePrice)]
set.seed(101)

elasticnet.model = cv.glmnet(X,y,type.measure = "mse",nfolds=10)

for (i in 0:10){
  assign(paste("elasticnet.model",i,sep = ""),
  cv.glmnet(X,y, type.measure = "mse",family="gaussian", alpha=i/10))
}

plot(elasticnet.model)
```
```{r}
test.X <- as.matrix(test)
pred<-predict(elasticnet.model, newx = test.X)[,1]
predictions<-exp(pred)
prediction5 <- data.frame(Id = Id_labels,SalePrice = predictions)
#prediction5
write.csv(prediction5, file = 'prediction5.csv', row.names = F)
```
*0.12445 from kaggle*

###Weighted model
```{r}
Weighted_model <- data.frame(Id = Id_labels, SalePrice = (prediction_ridge+3*prediction_lasso)/4)
write.csv(Weighted_model, file = 'prediction6.csv', row.names = F)
```
*0.12052 from kaggle*

##Generalized Additive Models

```{r}
saleprice <- data$SalePrice[!is.na(data$SalePrice)]
gam.lm.fit<-lm(saleprice ~ s(OverallQual,df=5) + s(Age,df=5) + s(TotalBsmtSF,df = 5) + s(LotArea,df= 5) ,data = train)
par(mfrow = c(2,2))
plot.Gam(gam.lm.fit, se = TRUE, col = 'darkolivegreen')
```
```{r}
pred<-exp(predict(gam.lm.fit,test))
prediction8<-data.frame(Id = Id_labels, SalePrice =  pred)
write.csv(prediction8, file = "prediction8.csv", row.names = FALSE)
```
*0.18084 from kaggle*

__Regression Trees__
```{r}
library(tree)
saleprice<-data$SalePrice[!is.na(data$SalePrice)]
train.tree<-data.frame(train)
tree.fit<- tree(saleprice~., data = train.tree)
summary(tree.fit)
par(mfrow = c(1,1))
plot(tree.fit)
text(tree.fit,pretty = 0)
set.seed(2019)
cv.out<-cv.tree(tree.fit, K = 10)
par(mfrow = c(1,2))
plot(cv.out$k,cv.out$dev,type = "b")
plot(cv.out$size,cv.out$dev,type = "b")
prune.out<-prune.tree(tree.fit,best = 8)
plot(prune.out)
text(prune.out,pretty= 0)
```

```{r}
test.x<-data.frame(test)
pred<-exp(predict(prune.out,newdata = test.x))
prediction9<-data.frame(Id = Id_labels, SalePrice =  pred)

write.csv(prediction9, file = "prediction9.csv", row.names = FALSE)
```
*0.21620 from Kaggle*

##Bagging
```{r}
bag.out<-randomForest(saleprice~.,data = train.tree, mtry = 178, importance =T)
summary(bag.out)
```
```{r}
importance(bag.out)
varImpPlot(bag.out)
test.x<-data.frame(test)
pred<-exp(predict(bag.out,newdata = test.x))
prediction10<-data.frame(Id = Id_labels, SalePrice =  pred)

write.csv(prediction10, file = "prediction10.csv", row.names = FALSE)
```
*0.13943 from Kaggle*

##Random Forest
```{r}
rf <- randomForest(saleprice ~., data = train.tree,mtry=round(sqrt(178)),importance = T)
rf
```
```{r}
importance(rf)
varImpPlot(rf)
test.x<-data.frame(test)
pred<-exp(predict(rf,newdata = test.x))
prediction11<-data.frame(Id = Id_labels, SalePrice =  pred)

write.csv(prediction11, file = "prediction11.csv", row.names = FALSE)
```
*0.14555 from Kaggle*

##Boosting
```{r}
gbm.cv.out <- gbm(saleprice ~., data = train.tree, distribution = "gaussian",shrinkage = 0.01, n.trees= 1000,interaction.depth = 4,cv.folds = 10)
which.min(gbm.cv.out$cv.error)
summary(gbm.cv.out)
```
```{r}
test.x<-data.frame(test)
pred<-exp(predict(gbm.cv.out,newdata = test.x))
prediction12<-data.frame(Id = Id_labels, SalePrice =  pred)
write.csv(prediction12, file = "prediction12.csv", row.names = FALSE)
```
*0.12553 from Kaggle*
